---
title: "Lab_2"
author: "Julia Stapledon"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", message = FALSE, warning = FALSE)
```

```{r import_data}
library(ggplot2movies)
data(movies)
library(dplyr)
library(knitr)
library(ggplot2)
```

Problem 1

-   What is the range of years of production of the movies in this dataset

```{r year_range}
cat("The range of years is from ", min(movies$year), " to ", max(movies$year))
```

Problem 2

-   What is the proportion of movies that have their budget included in this database, and what proportion doesn't? What are the top 5 most expensive movies in this dataset?

```{r movie_budgets}
included <- sum(!is.na(movies$budget))
omitted <- sum(is.na(movies$budget))
tot <- included + omitted
cat('Proportion of movies that have their budget listed: ', included/tot, "\n")
cat('Proportion of movies that do not have their budget listed: ', omitted/tot, "\n")
cat("Top 5 highest budget movies", "\n")
budgets <- select(arrange(movies, desc(budget)), title, budget)
kable(budgets[1:5,])
```

\newpage

Problem 3

-   What are the top 5 longest movies?

```{r longest_movies}
lengths <- select(arrange(movies, desc(length)), title, length)
kable(lengths[1:5,])
```

Problem 4

-   Of all short movies, which is the shortest (in minutes)? Which is the longest? How long are the shortest and the longest short movies?

```{r longest_short}
shorts <- filter(movies, Short == 1)
cat("The shortest short movie is: ",
    (head(arrange(shorts, length), n = 1))$title, " at ",
    min(shorts$length), "minutes in length\n")

cat("The longest short movie is: ",
    head(arrange(shorts, desc(length)), n = 1)$title, " at ",
    max(shorts$length), "minutes in length")
```

\newpage

Problem 5

-   How many movies of each genre are there in this database?

```{r genre_plot}
action <- nrow(filter(movies, Action == 1))
animation <- nrow(filter(movies, Animation == 1))
comedy <- nrow(filter(movies, Comedy == 1))
drama <- nrow(filter(movies, Drama == 1))
documentary <- nrow(filter(movies, Documentary == 1))
romance <- nrow(filter(movies, Romance == 1))
short <- nrow(filter(movies, Short == 1))
color <- c("deeppink","pink","lavender","purple","navy","blue","lightblue")
genre <- c("action", "animation", "comedy", "drama", "documentary", "romance", "short")

counts <- c(action, animation, comedy, drama, documentary, romance, short)

genres <- barplot(counts, main = "Movies by Genre",
                  xlab = "Genre", ylim = c(0, max(counts) + 5000), 
                  col = color, width = 20, space = 0.05)
legend("topleft", fill = color, legend = genre)
```

\newpage

Problem 6

-   What is the average rating of all movies within each genre?

```{r avg_rating_genre}
action <- filter(movies, Action == 1)$rating
avg_act <- mean(action, na.rm = T)
animation <- filter(movies, Animation == 1)$rating
avg_ani <- mean(animation, na.rm = T)
comedy <- filter(movies, Comedy == 1)$rating
avg_com <- mean(comedy, na.rm = T)
drama <- filter(movies, Drama == 1)$rating
avg_dra <- mean(drama, na.rm = T)
documentary <- filter(movies, Documentary == 1)$rating
avg_doc <- mean(documentary, na.rm = T)
romance <- filter(movies, Romance == 1)$rating
avg_rom <- mean(romance, na.rm = T)
short <- filter(movies, Short == 1)$rating
avg_sho <- mean(short, na.rm = T)
color <- c("deeppink","pink","lavender","purple","navy","blue","lightblue")
genre <- c("action", "animation", "comedy", "drama", "documentary", "romance", "short")
counts <- c(avg_act, avg_ani, avg_com, avg_doc, avg_dra, avg_rom, avg_sho)
par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
genres <- barplot(counts, main = "Average Rating of Movies by Genre",
                  xlab = "Genre", ylim = c(0, 10), 
                  col = color, width = 20, space = 0.05)
legend("right", inset=c(-0.33,0), fill = color, legend = genre)
```

\newpage

Problem 7

-   What is the average rating of all movies within each genre that were produced in the years 2000-2005?

```{r rating_genre_year}
action <- filter(filter(movies, Action == 1), year <= 2005 & year >= 2000)$rating
avg_ac <- mean(action, na.rm = T)
animation <- filter(filter(movies, Animation == 1), year <= 2005 & year >= 2000)$rating
avg_an <- mean(animation, na.rm = T)
comedy <- filter(filter(movies, Comedy == 1), year <= 2005 & year >= 2000)$rating
avg_co <- mean(comedy, na.rm = T)
drama <- filter(filter(movies, Drama == 1), year <= 2005 & year >= 2000)$rating
avg_dr <- mean(drama, na.rm = T)
documentary <- filter(filter(movies, Documentary == 1), year <= 2005 & year >= 2000)$rating
avg_do <- mean(documentary, na.rm = T)
romance <- filter(filter(movies, Romance == 1), year <= 2005 & year >= 2000)$rating
avg_ro <- mean(romance, na.rm = T)
short <- filter(filter(movies, Short == 1), year < 2005 & year > 2000)$rating
avg_sh <- mean(short, na.rm = T)
color <- c("deeppink","pink","lavender","purple","navy","blue","lightblue")
genre <- c("action", "animation", "comedy", "drama", "documentary", "romance", "short")
counts <- c(avg_ac, avg_an, avg_co, avg_do, avg_dr, avg_ro, avg_sh)
par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
genres <- barplot(counts, main = "Average Rating of Movies by Genre 2000-2005",
                  xlab = "Genre", ylim = c(0, 10), 
                  col = color, width = 20, space = 0.05)
legend("right", inset=c(-0.33,0), fill = color, legend = genre)
```

\newpage

Problem 8

-   For each of the first 6 genres (not including short movies) consider only movies from 1990 until the last year recorded and plot a function of the number of movies in this data base of corresponding genre produced by year, for years from 1990 until the last year recorded. For each of the 6 genres you should have one curve, and plot all the curves in the same figure.

```{r curve_movies}
df <- filter(movies, year >= 1990) %>%
group_by(year) %>%
summarise(act = sum(Action == 1),
          ani = sum(Animation == 1),
          com = sum(Comedy == 1),
          dra = sum(Drama == 1),
          doc = sum(Documentary == 1),
          rom = sum(Romance == 1))
ggplot(df, aes(x = year)) +
  geom_line(aes(y = act, color = "Action")) + 
  geom_line(aes(y = ani, color = "Animation")) + 
  geom_line(aes(y = com, color = "Comedy")) + 
  geom_line(aes(y = dra, color = "Drama")) +
  geom_line(aes(y = doc, color = "Documentary")) +
  geom_line(aes(y = rom, color = "Romance")) +
  xlab("Year") + ylab("Movies Made in Each Genre") +
  ggtitle("Number of Movies Made in Various Genres 1990-2005") +
  labs(colour= "Genres")
```

\newpage

Problem 9

-   plot each MPAA rating over the years 1990-2005

```{r mpaa}
df1 <- filter(movies, year > 1990 & !is.na(movies$mpaa)) %>%
  group_by(year) %>%
  summarise(R = sum(mpaa == "R"),
            PG13 = sum(mpaa == "PG-13"),
            PG = sum(mpaa == "PG"),
            NC17 = sum(mpaa == "NC-17"))
ggplot(df1, aes(x = year)) +
  geom_line(aes(y = R, color = "R")) +
  geom_line(aes(y = PG13, color = "PG-13")) +
  geom_line(aes(y = PG, color = "PG")) +
  geom_line(aes(y = NC17, color = "NC-17")) +
  xlab("Year") + ylab("Movies with Various MPAA ratings") +
  ggtitle("Number of Movies Made With Various MPAA Ratings 1990-2005") +
  labs(colour= "Rating")
```

\newpage

-   Plot the budget versus the rating given on IMDB

```{r budget_v_rating}
df = filter(movies, !is.na(movies$budget) & !is.na(movies$rating) )
data(df)
y = df$rating
x = df$budget

plot(x, y, pch = 16, xlab = "IMDB Rating",
     ylab = "Budget", main = "IMDB Rating vs. Budget",
     col = "deeppink")
```

\newpage

-   Plot votes versus rating

```{r votes_v_rating}
df = filter(movies, !is.na(movies$votes) & !is.na(movies$rating) )
data(df)
x = df$rating
y = df$votes

plot(x, y, pch = 16, xlab = "IMDB Rating",
     ylab = "Votes", main = "IMDB Rating vs. Votes",
     col = "deeppink")
```
