---
title: "lab_1"
author: "Julia Stapledon"
date: "2023-03-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r read_data}
df = read.csv("Absenteeism_at_work.csv", sep = ";", header = TRUE)
```

```{r height_vs_weight}
x = df$Weight
y = df$Height

plot(x, y,pch = 16, xlab = 'Weight', ylab  = 'Height',
     main = 'Height vs. Weight', col = "deeppink")
```

This graph doesn't provide much useful information. There is some grouping and some outliers but nothing meaningful for trends.

```{r hours_of_absences}
hist(df$Absenteeism.time.in.hours, xlab = "Hours", 
     main = "Hours of Absences", col = "deeppink")
```

Most people have less than 10 hours of being absent from work.

```{r age_vs_absences}
hist(df$Age, xlab = "Age",
     main = "Age Corresponding to Absences", col = "deeppink")
```

The most occurrences are between 35-40. There is not a curve but there is less occurrences on the right side of the histogram.

```{r hours_by_month}
counts <- table(df$Month.of.absence)
barplot(counts, main = "Absent Hours by Month", xlab = "Month", ylab = "Hours Absent",
         ylim = c(0 , max(counts) + 15), col = "deeppink")
```

there is not much to gather from this plot. The most absences were in March, but there are no trends. The 0 is people who had never been absent.

```{r social_smoker}
hours <- df$Absenteeism.time.in.hours
smoking_status <- df$Social.smoker
boxplot(hours~smoking_status, data = df, col = c("deeppink", "pink"), 
        main = "Social Smoking vs Hours Absent", horizontal = T)
legend("topright", legend = c("Non-Smoker", "Smoker"), fill = c("deeppink", "pink"))

```

There is no correlation between smoking and hours absent. 

``` {r social_drinker}
hours <- df$Absenteeism.time.in.hours
drinking_status <- df$Social.drinker
boxplot(hours~drinking_status, data = df, col = c("deeppink", "pink"),
        main = "Social Drinking vs. Hours", horizontal = T)
legend("topright", legend = c("Non-Drinker","Drinker"), fill = c("deeppink", "pink"))
```

Once again there isn't much to gather from this information. There are many outliers and this graph doesn't show much.